---
title: "Maps"
format: html
---

```{r}
#| include: FALSE
library(tidyverse);
library(mdsr);
library(maps);
library(viridis);
library(lubridate);
library(statebins);
library(leaflet);
library(sf);
library(htmltools);
library(glue);
library(rvest);
library(polite);
library(usdata);
```

```{r}
#| include: FALSE
states_sf <- 
  read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") |>
  select(name, geometry)

states_sf <- states_sf |>
  mutate(name = str_to_lower(str_squish(name))) |>
  filter(!(str_detect(name, "alaska|hawaii")))
```


```{r}
#| include: FALSE
states_map <- state_stats |> 
  mutate(region = tolower(state)) |>
  select(region, unempl, poverty) |>
  inner_join(map_data("state"))

int_states_map <- states_sf |> 
  inner_join(state_stats |> 
    mutate(region = tolower(state)) |>
    select(region, unempl, poverty),
  by =c("name" = "region"))
```

```{r}
#| include: FALSE
bins <- c(0, 2, 4, 6, 8, 10, 12, Inf)
pal <- colorBin("YlOrRd", domain = int_states_map, bins = bins)

int_states_map <- int_states_map |>
  mutate(labels = str_c(name, ": ", unempl, " % Unemployed"))

labels <- lapply(int_states_map$labels, HTML)
```

State-level graphs showing unemployment levels in 2010 by percentages
```{r}
#| echo: FALSE
ggplot(states_map, aes(map_id = region)) +
  geom_map(aes(fill = unempl), map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  scale_fill_viridis_c() +
  theme_void() +
  labs(x = "", y = "", fill = "Unemployment (%)")
```
Data aquired from the usdata package
```{r}
#| echo: FALSE
leaflet(int_states_map) |>
  setView(-96, 37.8, 4) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(unempl),
    weight = 3,
    opacity = 1,
    color = "white",
    dashArray = "4",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#000",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "bold", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) |>
  addLegend(pal = pal, values = ~unempl, opacity = 0.7, title = NULL,
    position = "bottomright")
```
Data aquired from the usdata package

From these graphs we can see that there is a national average of around 4-8% unemployed with some notable outliers California, Nevada, and Rhode island all with unemployment over 10%. 

```{r}
#| include: FALSE
bins <- c(0, 4, 8, 12, 16, 20, Inf)
pal <- colorBin("YlOrRd", domain = int_states_map, bins = bins)

int_states_map <- int_states_map |>
  mutate(labels = str_c(name, ": ", poverty, " % under the poverty line"))

labels <- lapply(int_states_map$labels, HTML)
```

Interestingly, this does not correlate with national poverty levels. Nevada and California are average in national poverty while Mississippi is a significant outlier despite have an around average unemployment rate. 

State-level graphs showing poverty levels in 2010 by percentages
```{r}
#| echo: FALSE
ggplot(states_map, aes(map_id = region)) +
  geom_map(aes(fill = poverty), map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  scale_fill_viridis_c() +
  theme_void() +
  labs(x = "", y = "", fill = "Poverty (%)")
```
Data aquired from the usdata package
```{r}
#| echo: FALSE
leaflet(int_states_map) |>
  setView(-96, 37.8, 4) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(poverty),
    weight = 3,
    opacity = 1,
    color = "white",
    dashArray = "4",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#000",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "bold", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) |>
  addLegend(pal = pal, values = ~poverty, opacity = 0.7, title = NULL,
    position = "bottomright")
```
Data aquired from the usdata package