---
title: "Project 1"
format: html
---

```{r}
#| include: FALSE
library(tidyverse);
library(RTextTools);
library(tidytext);
library(textdata);
library(wordcloud);
library(wordcloud2);
library(viridis);
library(ggthemes);
library(stopwords);
library(chameleon);
data(NYTimes);
as_tibble(NYTimes);
```

The NYTimes dataset contains text of all the titles for articles from 1996 - 2006. I want to look at how the sentiment, topics, and word choice changes over the years.

```{r}
#| include: FALSE
months <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
```


```{r}
# Bing sentiments for titles separated by month
NYTimes |>
  mutate (
    # Extract the month and convert to a number, regex 1, str_extract
    month = match(str_extract(Date, "[A-z]{3}"), months), 
    year = str_extract(Date, "[0-9]{2}$"), # Extract the date, first instance of 2 digits, regex 2
    # Append the first digits of the year (19/20) onto the corresponding years, str_detect, str_c
    year = ifelse(str_detect(year, "^9"), str_c("19", year), str_c("20", year)),
    yearMonth = str_c(year, "-", month), 
    title = as.character(Title) # unnest_tokens requires data in a character vector
  ) |>
  unnest_tokens(word, title, token = "words") |> # Extract all the words
  inner_join(get_sentiments(lexicon = "bing")) |> # Keep only words in the sentiment dataset
  group_by(year, month, yearMonth) |>
  mutate (
    sentiment = sentiment == "negative" # Convert to boolean
  ) |>
  summarize(sentimentScore = mean(sentiment) ) |> # Text analysis 1
  arrange(yearMonth) |>
  ggplot(aes(x = month,  y = sentimentScore)) +
  geom_col() +
  scale_x_continuous(breaks = seq(1:12)) + # Set the x-axis labels
  labs(x = "Month", y = "% Negative Sentiment") +
  theme_minimal() +
  facet_wrap(scales = "fixed", ~year)
```

Most months have an overall negative sentiment score (greater than %50). There are varying number of articles per month so some months only have a few data points while others have a lot more. Some notable points are May 1997, February 1998, November 2002 and 2003 all of which have 100% negative sentiment. These coincide with natural disasters and terrorist attacks being the only reported on news in the data base.

```{r}
#| include: FALSE
codes <- c("Macroeconomics", "Civil Rights", "Health", "Agriculture", "Labor, Immigration, and Employment", "Education", "Environment", "Energy", "Immigration", "Transportation", "", "Law", "Social Welfare", "Community Development and Housing", "Finance", "Defense", "Science", "Foreign Trade", "International Affairs", "Federal Government", "Public Lands and Water Management", "", "Arts and Entertainment", "State and Local Government Administration", "", "Weather and Natural Disasters", "Fires", "", "Sports and Recreation", "Death Notices", "Churches and Religion", "Other, Miscellaneous, and Human Interest")
```


```{r}
# Top four topics for each year
NYTimes |>
  mutate(
    year = str_extract(Date, "[0-9]{2}$"), # Extract the date, first instance of 2 digits
    # Append the first digits of the year (19/20) onto the corresponding years
    year = ifelse(str_detect(year, "^9"), str_c("19", year), str_c("20", year))
  ) |>
  group_by(year) |>
  count(Topic.Code)|>
  mutate(
    topic = codes[Topic.Code] # Use the topic code as an index to convert to the actual topic
  ) |>
  arrange(desc(n)) |>
  slice(1:4) |> # Keep only the 4 most occurent topics for each
  ggplot(aes(x = topic, y = n, fill = topic)) +
  geom_col() +
  labs(x = "", y = "Count", fill = "Topic") +
  theme_classic() +
  theme(
    axis.text.x = element_blank(), 
    axis.line = element_line(linewidth = 0.1)
    ) +
  facet_wrap(scales = "free", ~year)
```

The most common topics change over the years. Up until 2001 the Federal Government was a large part of the news while Defense was a much smaller part. After and during 2001 Defense because a much larger part of the news. This coincides with 9/11 and the war in Iraq. The dataset ends at 2006 which is still during the war in Iraq so Defense is still a large part of the news. International Affairs has always been a large part of the news which makes sense as the NYT is not a solely domestic news source.

```{r, fig.width=10,fig.height=7}

# Top 10 most common words for each year
NYTimes |>
  mutate (
    year = str_extract(Date, "[0-9]{2}$"),
    # Append the first digits of the year (19/20) onto the corresponding years
    year = ifelse(str_detect(year, "^9"), str_c("19", year), str_c("20", year)),
    title = as.character(Title) # unnest_tokens requires data in a character vector
  ) |>
  unnest_tokens(word, title, token = "words") |> # Separate all words
  anti_join(get_stopwords(source = "smart")) |> # Remove all common words
  filter(!str_detect(word, "u.s|^[0-9]+$")) |> # Filter numbers and mentions of u.s, regex 3
  group_by(year) |>
  count(word, sort = TRUE) |> # Text analysis 2
  slice(1:10) |>
  ggplot(aes(fct_reorder(word, n), n)) +
  geom_col() +
  coord_flip() +
  labs(x = "Count", y = "Word") +
  theme_minimal() +
  facet_wrap(scales = "free", ~year)
```

Up until 2001 a majority of the news was focused on politics with words like "campaign" and "politics" as well as names like "clinton" and "dole" being the most commonly used words. Same as before, after 2001 Iraq becomes a huge part of the news. Words like "Iraq" and "war" become huge parts of the news and politics becomes a lesser part of the news.
